---
title: "Final EGRET v1 Banovich application - Data Analysis"
author: "Deborah Weighill"
date: "July 5, 2020"
output:
  html_notebook:
    toc: true
    theme: cosmo
    df_print: paged
---

# Overview
Here we do some analysis on the banovich results for the "application". We start from the "master difference table" that we created.



# Load libraries
```{r}
library(dplyr)
library(ggplot2)
library(ggpubr)
library(reshape2)
library(jcolors)
library(gridExtra)
library(grid)
library(gridExtra)
```

# Load difference tables
```{r}
load("../outputs/tf_diff_degree_table_banovich_smart1_07042020.RData")
load("../outputs/gene_diff_degree_table_banovich_smart1_7042020.RData")

```

# set color pallet
```{r}
#pallet2 <- c("#42858C", "#E48F1B","#570D32")
#pallet2 <- c("#EB8F1E","#AD1F2F","#2592A7")
#pallet2 <- c("#0BB19F", "#FF4500","#570D32")    
#pallet2 <- c("#0BB19F", "#FF4500","#50D32")  
pallet2 <- c(    "#FFBF00",  "#870E75"  ,  "#0BB19F" )
```


We can now plot distributions of tf degree and gene degree in difference network for each tissue.
```{r}
colnames(tf_degrees)[1] <- "tf"
tf_degrees_scaled <- scale(tf_degrees[,c(2:361)], center = TRUE, scale = TRUE)
tf_degrees_scaled <- cbind(tf_degrees[,c(1)], tf_degrees_scaled, tf_degrees[,c(362:367)])
colnames(tf_degrees_scaled)[1] <-"tf"
tf_degree_lcl <- melt(tf_degrees[,c(1,c(3:121))], id.vars = c("tf"))
tf_degree_lcl$tissue <- "LCL"
tf_degree_iPSC <- melt(tf_degrees[,c(1,c(123:241))], id.vars = c("tf"))
tf_degree_iPSC$tissue <- "iPSC"
tf_degree_iPSC_CM <- melt(tf_degrees[,c(1,c(243:361))], id.vars = c("tf"))
tf_degree_iPSC_CM$tissue <- "iPSC-CM"
tf_degrees_melted <- rbind(tf_degree_lcl,tf_degree_iPSC,tf_degree_iPSC_CM)
colnames(tf_degrees_melted) <- c("tf","individual","diff_out_degree","Tissue")

pdf("banovich_tf_disruption_boxplot.pdf",width = 4, height = 3.5)
p <- ggplot(tf_degrees_melted, aes(x=Tissue, y=diff_out_degree, color=Tissue)) + geom_boxplot() + theme_bw() + scale_color_manual(values=pallet2) + labs(title="Distribution of TF Disruption Scores", x ="Cell type", y = "TF Disruption Score") + theme(legend.position = "none")
plot(p)
dev.off()


ggplot(tf_degrees_melted[which(tf_degrees_melted$Tissue == "iPSC"),], aes(x=Tissue, y=diff_out_degree, color=Tissue)) + geom_violin() + theme_linedraw() + scale_color_manual(values=c("red")) + labs(title="Distribution of TF Disruption Scores", x = NULL, y = "TF Disruption Score") + theme(legend.position = "right")

ggplot(tf_degrees_melted[which(tf_degrees_melted$Tissue == "iPSC-CM"),], aes(x=Tissue, y=diff_out_degree, color=Tissue)) + geom_violin() + theme_linedraw() + scale_color_manual(values=c("blue")) + labs(title="Distribution of TF Disruption Scores", x =NULL, y = "TF Disruption Score") + theme(legend.position = "right")

ggplot(tf_degrees_melted[which(tf_degrees_melted$Tissue == "LCL"),], aes(x=Tissue, y=diff_out_degree, color=Tissue)) + geom_violin() + theme_linedraw() + scale_color_manual(values=c("orange")) + labs(title="Distribution of TF Disruption Scores", x =NULL, y = "TF Disruption Score") + theme(legend.position = "right")

ggplot(tf_degrees_melted[which(tf_degrees_melted$individual == "iPSC-CM_indiv_21"),], aes(x=Tissue, y=diff_out_degree, color=Tissue)) + geom_violin() + theme_linedraw() + scale_color_manual(values=c("blue")) + labs(title="Distribution of TF Disruption Scores", x =NULL, y = "TF Disruption Score") + theme(legend.position = "right")

```

scaled plots
```{r}

tf_degree_lcl <- melt(tf_degrees_scaled[,c(1,c(3:121))], id.vars = c("tf"))
tf_degree_lcl$tissue <- "LCL"
tf_degree_iPSC <- melt(tf_degrees_scaled[,c(1,c(123:241))], id.vars = c("tf"))
tf_degree_iPSC$tissue <- "iPSC"
tf_degree_iPSC_CM <- melt(tf_degrees_scaled[,c(1,c(243:361))], id.vars = c("tf"))
tf_degree_iPSC_CM$tissue <- "iPSC-CM"
tf_degrees_melted <- rbind(tf_degree_lcl,tf_degree_iPSC,tf_degree_iPSC_CM)
colnames(tf_degrees_melted) <- c("tf","individual","diff_out_degree","Tissue")

pdf("banovich_scaled_tf_disruption_boxplot.pdf",width = 4, height = 3.5)
p <- ggplot(tf_degrees_melted, aes(x=Tissue, y=diff_out_degree, color=Tissue)) + geom_boxplot() + theme_bw() + scale_color_manual(values=pallet2) + labs(title="Distribution of scaled TF disruption scores", x ="Cell type", y = "Scaled TF disruption score") + theme(legend.position="none")
plot(p)
dev.off()

pdf("banovich_scaled_tf_disruption_violin.pdf",width = 4, height = 3.5)
ggplot(tf_degrees_melted, aes(x=Tissue, y=diff_out_degree, fill=Tissue, col = Tissue)) + geom_violin() + theme_bw() + scale_fill_manual(values=pallet2) + scale_color_manual(values=pallet2) + labs(title="Distribution of TF Disruption Scores", x ="Cell type", y = "TF Disruption Score") + theme(legend.position="none")
dev.off()

ggplot(tf_degrees_melted, aes(x=Tissue, y=diff_out_degree, color=Tissue,)) + geom_violin() + theme_bw() + scale_color_manual(values=pallet2) + labs(title="Distribution of scaled TF disruption scores", x ="Cell type", y = "Scaled TF disruption score") +  ylim(-2, 2) + theme(legend.position="none")

p1 <- ggplot(tf_degrees_melted, aes(x=diff_out_degree,  color=Tissue)) + geom_density() + theme_bw() + scale_color_manual(values=pallet2) +labs(title=NULL, x ="Scaled TF disruption score", y = "Frequency") 

p2 <- ggplot(tf_degrees_melted, aes(x=diff_out_degree,  color=Tissue)) + geom_density() + theme_bw() + scale_color_manual(values=pallet2) +labs(title=NULL, x ="Scaled TF disruption score", y = "Frequency") +  xlim(-1, 1)

p3 <- grid.arrange(p1 + theme(legend.position="hidden"), p2 + theme(legend.position="right"), nrow=1, widths = c(3,4))
p4 <- annotate_figure(p3,top = text_grob("Distribution of scaled TF disruption scores", color = "black", face = "bold", size = 14))
pdf("disruption_score_dist.pdf",width = 7, height = 3)
plot(p4)
dev.off()


```

```{r}
gene_degrees_scaled <- scale(gene_degrees[,c(2:361)], center = TRUE, scale = TRUE)
gene_degrees_scaled <- cbind(gene_degrees[,c(1)], gene_degrees_scaled, gene_degrees[,c(362:367)])
colnames(gene_degrees_scaled)[1] <- "gene"
gene_degree_lcl <- melt(gene_degrees_scaled[,c(1,c(3,121))], id.vars = c("gene"))
gene_degree_lcl$tissue <- "LCL"
gene_degree_iPSC <- melt(gene_degrees_scaled[,c(1,c(123,241))], id.vars = c("gene"))
gene_degree_iPSC$tissue <- "iPSC"
gene_degree_iPSC_CM <- melt(gene_degrees_scaled[,c(1,c(243,361))], id.vars = c("gene"))
gene_degree_iPSC_CM$tissue <- "iPSC-CM"
gene_degrees_melted <- rbind(gene_degree_lcl,gene_degree_iPSC,gene_degree_iPSC_CM)
colnames(gene_degrees_melted) <- c("gene","individual","diff_in_degree","tissue")
ggplot(gene_degrees_melted, aes(x=tissue, y=diff_in_degree, fill=tissue, col = tissue)) + geom_boxplot() + theme_linedraw() + scale_fill_manual(values=pallet2) + scale_color_manual(values=pallet2) + labs(title="Distribution of Gene Disruption Scores", x ="Cell type", y = "Gene Disruption Score")

```






# Enrichment tests on TFs
Let's try and see, for a given individual, are "high difference" TFs enriched in GWAS hits.

## CD
```{r}
d = 20
individuals <- c(c(3:121),c(123:241),c(243:361))
CD_enrich <- sapply(individuals, FUN = function(i){
  high_diff_gwas <- length(which((tf_degrees[,i] >d) & (tf_degrees$GWAS_CD_tf == 1)))
  low_diff_gwas <- length(which((tf_degrees[,i] <=d) & (tf_degrees$GWAS_CD_tf == 1)))
  high_diff_notGwas <- length(which((tf_degrees[,i] >d) & (tf_degrees$GWAS_CD_tf == 0)))
  low_diff_notGwas <- length(which((tf_degrees[,i] <=d) & (tf_degrees$GWAS_CD_tf == 0)))
  contingency_table <- matrix(c(high_diff_gwas,low_diff_gwas,high_diff_notGwas,low_diff_notGwas), nrow = 2,dimnames = list(Diff = c("High", "Low"),GWAS = c("yes", "no")))
  test <- fisher.test(contingency_table, alternative = "greater")
  return(as.numeric(test$p.value))
})
names(CD_enrich) <- colnames(tf_degrees)[individuals]
cd_enriched_indiv <- names(which(CD_enrich < 0.05))
cd_enriched_indiv

cd_df <- as.data.frame(CD_enrich)
colnames(cd_df) <- "pvalue"
cd_df$Tissue <- NA
cd_df$Tissue[1:119] <- "LCL"
cd_df$Tissue[120:238] <- "iPSC"
cd_df$Tissue[239:357] <- "iPSC-CM"
cd_df$Individual <- NA
cd_df$Individual <- c(c(1:119),c(1:119),c(1:119))
#pallet2 <- c("#FFA733","#2CAFC9","#AD5C99")


#pallet2 <- c("#42858C", "#E48F1B","#570D32")
#pallet2 <- c("#EB8F1E","#AD1F2F","#2592A7")
#pallet2 <- c("#0BB19F", "#FF4500","#570D32")    

pdf("banovich_CD_enrichment.pdf",width = 5, height = 3.5)
ggplot(cd_df) + geom_point(shape = 19, aes(x = reorder(Individual, log10(pvalue)), y = -log10(pvalue), col = Tissue),) + theme_classic()+ scale_color_manual(values=pallet2) + geom_abline(slope = 0, intercept = -log10(0.05), linetype = "dashed", col = "black") + xlab("Individual") + theme( axis.text.x=element_blank(), axis.ticks.x=element_blank()) + labs(y = "-log10(p-value)", title = "Enrichment of disrupted TFs (d > 15) in CD genes",color = "Cell type") + annotate("text", x = 110, y = 1.4, label = "p = 0.05")
dev.off()
```
```{r}
tf_degrees[,c("tf",cd_enriched_indiv)]
tf_degrees[which(tf_degrees$GWAS_CD_tf == 1),c("tf","GWAS_CD_tf",cd_enriched_indiv)]
cd_tf_table <- diff_table[(which(diff_table$tf == c("ELF1","FOXP1,FOXP2"))),c("tf","gene",cd_enriched_indiv)]
cd_tf_table[apply(cd_tf_table[,cd_enriched_indiv], 1, function(x) any(x > 0.5)), ]
```
the first gene ENSG00000110768 is maybe differentially expressed in chrons disease
"Pathogenicity of In Vivo Generated Intestinal Th17 Lymphocytes is IFNγ Dependent"

maybe look here for the second gene (ENSG00000167748) 
Genetically-modified stem cells in treatment of human diseases: Tissue kallikrein (KLK1)-based targeted therapy (Review)

## CD
```{r}
d = 2
individuals <- c(c(3:121),c(123:241),c(243:361))
CD_enrich <- sapply(individuals, FUN = function(i){
  high_diff_gwas <- length(which((tf_degrees_scaled[,i] >d) & (tf_degrees_scaled$GWAS_CD_tf == 1)))
  low_diff_gwas <- length(which((tf_degrees_scaled[,i] <=d) & (tf_degrees_scaled$GWAS_CD_tf == 1)))
  high_diff_notGwas <- length(which((tf_degrees_scaled[,i] >d) & (tf_degrees_scaled$GWAS_CD_tf == 0)))
  low_diff_notGwas <- length(which((tf_degrees_scaled[,i] <=d) & (tf_degrees_scaled$GWAS_CD_tf == 0)))
  contingency_table <- matrix(c(high_diff_gwas,low_diff_gwas,high_diff_notGwas,low_diff_notGwas), nrow = 2,dimnames = list(Diff = c("High", "Low"),GWAS = c("yes", "no")))
  test <- fisher.test(contingency_table, alternative = "greater")
  return(as.numeric(test$p.value))
})
names(CD_enrich) <- colnames(tf_degrees_scaled)[individuals]
cd_enriched_indiv <- names(which(CD_enrich < 0.05))
cd_enriched_indiv

cd_df <- as.data.frame(CD_enrich)
colnames(cd_df) <- "pvalue"
cd_df$Tissue <- NA
cd_df$Tissue[1:119] <- "LCL"
cd_df$Tissue[120:238] <- "iPSC"
cd_df$Tissue[239:357] <- "iPSC-CM"
cd_df$Individual <- NA
cd_df$Individual <- c(c(1:119),c(1:119),c(1:119))

pdf("banovich_CD_enrichment_scaled.pdf",width = 5, height = 3.5)
ggplot(cd_df) + geom_point(shape = 19, aes(x = reorder(Individual, log10(pvalue)), y = -log10(pvalue), col = Tissue),) + theme_classic()+ scale_color_manual(values=pallet2) + geom_abline(slope = 0, intercept = -log10(0.05), linetype = "dashed", col = "black") + xlab("Individual") + theme( axis.text.x=element_blank(), axis.ticks.x=element_blank()) + labs(y = "-log10(p-value)", title = "Enrichment of disrupted TFs (d' > 2) in CD genes",color = "Cell type") + annotate("text", x = 110, y = 1.4, label = "p = 0.05")
dev.off()
```

```{r}
cd_enriched_indiv
tf_degrees_scaled[which(tf_degrees_scaled$GWAS_CD_tf == 1),c("tf","GWAS_CD_tf",cd_enriched_indiv)]

```


## CAD
```{r}
d = 15
individuals <- c(c(3:121),c(123:241),c(243:361))
CAD_enrich <- sapply(individuals, FUN = function(i){
  high_diff_gwas <- length(which((tf_degrees[,i] >d) & (tf_degrees$GWAS_CAD_tf == 1)))
  low_diff_gwas <- length(which((tf_degrees[,i] <=d) & (tf_degrees$GWAS_CAD_tf == 1)))
  high_diff_notGwas <- length(which((tf_degrees[,i] >d) & (tf_degrees$GWAS_CAD_tf == 0)))
  low_diff_notGwas <- length(which((tf_degrees[,i] <=d) & (tf_degrees$GWAS_CAD_tf == 0)))
  contingency_table <- matrix(c(high_diff_gwas,low_diff_gwas,high_diff_notGwas,low_diff_notGwas), nrow = 2,dimnames = list(Diff = c("High", "Low"),GWAS = c("yes", "no")))
  test <- fisher.test(contingency_table, alternative = "greater")
  return(as.numeric(test$p.value))
})
names(CAD_enrich) <- colnames(tf_degrees)[individuals]

cad_enriched_indiv <- names(which(CAD_enrich < 0.05))
cad_enriched_indiv
cad_df <- as.data.frame(CAD_enrich)
colnames(cad_df) <- "pvalue"
cad_df$Tissue <- NA
cad_df$Tissue[1:119] <- "LCL"
cad_df$Tissue[120:238] <- "iPSC"
cad_df$Tissue[239:357] <- "iPSC-CM"
cad_df$Individual <- NA
cad_df$Individual <- c(c(1:119),c(1:119),c(1:119))

pdf("banovich_CAD_enrichment.pdf",width = 5, height = 3.5)
ggplot(cad_df) + geom_point(shape = 19, aes(x = reorder(Individual, log10(pvalue)), y = -log10(pvalue), col = Tissue),) + theme_classic()+ scale_color_manual(values=pallet2) + geom_abline(slope = 0, intercept = -log10(0.05), linetype = "dashed", col = "black") + xlab("Individual") + theme( axis.text.x=element_blank(), axis.ticks.x=element_blank())  + labs(y = "-log10(p-value)", title = "Enrichment of disrupted TFs (d > 15) in CAD genes",color = "Cell type") + annotate("text", x = 110, y = 1.4, label = "p = 0.05")
dev.off()
```

## CAD
```{r}
d = 2
individuals <- c(c(3:121),c(123:241),c(243:361))
CAD_enrich <- sapply(individuals, FUN = function(i){
  high_diff_gwas <- length(which((tf_degrees_scaled[,i] >d) & (tf_degrees_scaled$GWAS_CAD_tf == 1)))
  low_diff_gwas <- length(which((tf_degrees_scaled[,i] <=d) & (tf_degrees_scaled$GWAS_CAD_tf == 1)))
  high_diff_notGwas <- length(which((tf_degrees_scaled[,i] >d) & (tf_degrees_scaled$GWAS_CAD_tf == 0)))
  low_diff_notGwas <- length(which((tf_degrees_scaled[,i] <=d) & (tf_degrees_scaled$GWAS_CAD_tf == 0)))
  contingency_table <- matrix(c(high_diff_gwas,low_diff_gwas,high_diff_notGwas,low_diff_notGwas), nrow = 2,dimnames = list(Diff = c("High", "Low"),GWAS = c("yes", "no")))
  test <- fisher.test(contingency_table, alternative = "greater")
  return(as.numeric(test$p.value))
})
names(CAD_enrich) <- colnames(tf_degrees_scaled)[individuals]

cad_enriched_indiv <- names(which(CAD_enrich < 0.05))
cad_enriched_indiv
cad_df <- as.data.frame(CAD_enrich)
colnames(cad_df) <- "pvalue"
cad_df$Tissue <- NA
cad_df$Tissue[1:119] <- "LCL"
cad_df$Tissue[120:238] <- "iPSC"
cad_df$Tissue[239:357] <- "iPSC-CM"
cad_df$Individual <- NA
cad_df$Individual <- c(c(1:119),c(1:119),c(1:119))

pdf("banovich_CAD_enrichment_scaled.pdf",width = 5, height = 3.5)
ggplot(cad_df) + geom_point(shape = 19, aes(x = reorder(Individual, log10(pvalue)), y = -log10(pvalue), col = Tissue),) + theme_classic()+ scale_color_manual(values=pallet2) + geom_abline(slope = 0, intercept = -log10(0.05), linetype = "dashed", col = "black") + xlab("Individual") + theme( axis.text.x=element_blank(), axis.ticks.x=element_blank()) + labs(y = "-log10(p-value)", title = "Enrichment of disrupted TFs (d' > 2) in CAD genes",color = "Cell type") + annotate("text", x = 110, y = 1.4, label = "p = 0.05")
dev.off()
```






# Some t tests and box plots

```{r}
library(ggpubr)
library(dplyr)

tf_degree_lcl <- melt(tf_degrees_scaled[,c(1,c(3:121))], id.vars = c("tf"))
tf_degree_lcl$tissue <- "LCL"
tf_degree_iPSC <- melt(tf_degrees_scaled[,c(1,c(123:241))], id.vars = c("tf"))
tf_degree_iPSC$tissue <- "iPSC"
tf_degree_iPSC_CM <- melt(tf_degrees_scaled[,c(1,c(243:361))], id.vars = c("tf"))
tf_degree_iPSC_CM$tissue <- "iPSC-CM"
tf_degrees_melted <- rbind(tf_degree_lcl,tf_degree_iPSC,tf_degree_iPSC_CM)
colnames(tf_degrees_melted) <- c("tf","individual","diff_out_degree","tissue")
ggplot(tf_degrees_melted, aes(x=tissue, y=diff_out_degree, color=tissue)) + geom_violin()

```

```{r}
colnames(tf_degree_iPSC_CM) <-  c("tf","individual","diff_out_degree","tissue")
tf_degree_iPSC_CM$CAD <- tf_degrees_scaled$GWAS_CAD_tf[match(tf_degree_iPSC_CM$tf, tf_degrees_scaled$tf)]

test <- t.test(tf_degree_iPSC_CM[which(tf_degree_iPSC_CM$CAD == 1),]$diff_out_degree,tf_degree_iPSC_CM[which(tf_degree_iPSC_CM$CAD == 0),]$diff_out_degree, alternative="greater")

mean0 <- mean(tf_degree_iPSC_CM[which(tf_degree_iPSC_CM$CAD == 0),]$diff_out_degree)
mean1 <- mean(tf_degree_iPSC_CM[which(tf_degree_iPSC_CM$CAD == 1),]$diff_out_degree)

p1 <- ggplot(tf_degree_iPSC_CM, aes(x = as.factor(CAD), y = diff_out_degree, group = as.factor(CAD), fill = as.factor(CAD))) + scale_fill_manual(values = pallet2) + geom_violin() + theme_bw() + geom_point(x = as.factor(0), y = mean0, shape=18,size=5, color="#E41A1C")+ geom_point(x = as.factor(1), y = mean1, shape=18,size=5, color="#E41A1C") + labs(title = "Distribution of disruption scores for CAD and non-CAD TFs in iPSC-CMs",y = "TF disruption score", x = "CAD TF?") + theme(legend.position = "none")

p2 <- ggplot(tf_degree_iPSC_CM, aes(x = as.factor(CAD), y = diff_out_degree, group = as.factor(CAD), fill = as.factor(CAD))) + scale_fill_manual(values = pallet2) + geom_violin() + theme_bw() + geom_point(x = as.factor(0), y = mean0, shape=18,size=5, color="#E41A1C")+ geom_point(x = as.factor(1), y = mean1, shape=18,size=5, color="#E41A1C")+ ylim(-1,1) + labs(title = "Distribution of disruption scores for CAD and non-CAD TFs in iPSC-CMs",y = "TF disruption score", x = "CAD TF?") + theme(legend.position = "none")

p3 <- grid.arrange(p1 + theme(legend.position="none"), p2 + theme(legend.position="none"), nrow=1, widths = c(3,3))
pdf("cad_1_0_ipsc_cm.pdf",width = 6, height = 3)
plot(p3)
dev.off()

ggplot(tf_degree_iPSC_CM, aes(x = as.factor(CAD), y = diff_out_degree,group = as.factor(CAD))) + geom_jitter(position=position_jitter(0.2), size = 0.1) + theme_bw() + ylim(-1,1) + geom_point(aes(x = as.factor(0), y = mean0), shape=18,size=5, color="#E41A1C")+ geom_point(x = as.factor(1), y = mean1, shape=18,size=5, color="#E41A1C")


colnames(tf_degree_iPSC_CM) <-  c("tf","individual","diff_out_degree","tissue")
tf_degree_iPSC_CM$CD <- tf_degrees_scaled$GWAS_CD_tf[match(tf_degree_iPSC_CM$tf, tf_degrees_scaled$tf)]

t.test(tf_degree_iPSC_CM[which(tf_degree_iPSC_CM$CD == 1),]$diff_out_degree,tf_degree_iPSC_CM[which(tf_degree_iPSC_CM$CD == 0),]$diff_out_degree, alternative="greater")




colnames(tf_degree_lcl) <-  c("tf","individual","diff_out_degree","tissue")
tf_degree_lcl$CD <- tf_degrees_scaled$GWAS_CD_tf[match(tf_degree_lcl$tf, tf_degrees_scaled$tf)]

t.test(tf_degree_lcl[which(tf_degree_lcl$CD == 1),]$diff_out_degree,tf_degree_lcl[which(tf_degree_lcl$CD == 0),]$diff_out_degree, alternative="greater")


colnames(tf_degree_lcl) <-  c("tf","individual","diff_out_degree","tissue")
tf_degree_lcl$CAD <- tf_degrees_scaled$GWAS_CAD_tf[match(tf_degree_lcl$tf, tf_degrees_scaled$tf)]
t.test(tf_degree_lcl[which(tf_degree_lcl$CAD == 1),]$diff_out_degree,tf_degree_lcl[which(tf_degree_lcl$CAD == 0),]$diff_out_degree, alternative="greater")


```
# Testing for pallet design

```{r}
d = 2
individuals <- c(c(3:121),c(123:241),c(243:361))
CD_enrich <- sapply(individuals, FUN = function(i){
  high_diff_gwas <- length(which((tf_degrees_scaled[,i] >d) & (tf_degrees_scaled$GWAS_CD_tf == 1)))
  low_diff_gwas <- length(which((tf_degrees_scaled[,i] <=d) & (tf_degrees_scaled$GWAS_CD_tf == 1)))
  high_diff_notGwas <- length(which((tf_degrees_scaled[,i] >d) & (tf_degrees_scaled$GWAS_CD_tf == 0)))
  low_diff_notGwas <- length(which((tf_degrees_scaled[,i] <=d) & (tf_degrees_scaled$GWAS_CD_tf == 0)))
  contingency_table <- matrix(c(high_diff_gwas,low_diff_gwas,high_diff_notGwas,low_diff_notGwas), nrow = 2,dimnames = list(Diff = c("High", "Low"),GWAS = c("yes", "no")))
  test <- fisher.test(contingency_table, alternative = "greater")
  return(as.numeric(test$p.value))
})
names(CD_enrich) <- colnames(tf_degrees_scaled)[individuals]
cd_enriched_indiv <- names(which(CD_enrich < 0.05))
cd_enriched_indiv
cd_df <- as.data.frame(CD_enrich)
colnames(cd_df) <- "pvalue"
cd_df$Tissue <- NA
cd_df$Tissue[1:119] <- "LCL"
cd_df$Tissue[120:238] <- "iPSC"
cd_df$Tissue[239:357] <- "iPSC-CM"
cd_df$Individual <- NA
cd_df$Individual <- c(c(1:119),c(1:119),c(1:119))

 # blue_yonder   arctic_lime    mardi_gras safety_orange  tiffany_blue 
 #   "#3E71A8"     "#CEFF1A"     "#870E75"     "#FE6900"     "#0BB19F" 
#   kelly_green rich_electric_blue        maximum_red     majorelle_blue 
 #        "#29BF12"          "#00A5CF"          "#DE1A1A"          "#574AE2" 
#fluorescent_orange 
     #    "#FFBF00" 
    
pallet2 <- c(    "#FFBF00",  "#870E75"  ,  "#0BB19F" )

ggplot(cd_df) + geom_point(shape = 19, aes(x = reorder(Individual, log10(pvalue)), y = -log10(pvalue), col = Tissue),) + theme_classic()+ scale_color_manual(values=pallet2) + geom_abline(slope = 0, intercept = -log10(0.05), linetype = "dashed", col = "black") + xlab("Individual") + theme( axis.text.x=element_blank(), axis.ticks.x=element_blank()) + labs(y = "-log10(p-value)", title = "Enrichment of disrupted TFs (d' > 2) in CD genes",color = "Cell type") + annotate("text", x = 110, y = 1.4, label = "p = 0.05")
```



